Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}

### 2022-02-21 Paper.tex
Fix gas specification in CALL* operations
Fix the upsidedown multiplier in the bloom filter footnote

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \hyperlink{bigvee}{\bigvee}_{x \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(x) \big)
\end{equation}
where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence. It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
\begin{eqnarray}
M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{2047 - m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
\end{eqnarray}
where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
Notably, it treats $\mathbf{x}$ as big-endian (more significant bits will have smaller indices).

@@

&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A', z, \mathbf{o}) \equiv \begin{cases}

@@

\big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, A, 0, () \big) & \text{otherwise} \end{cases}$ \\

@@

@@

(\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\

@@

&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\

@@

&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\

@@

&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\

>>>>

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \hyperlink{bigvee}{\bigvee}_{x \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(x) \big)
\end{equation}
where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence. It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{2048 $= 2^{11}$(11 bits), and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
\begin{eqnarray}
M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{2047 - m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
\end{eqnarray}
where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
Notably, it treats $\mathbf{x}$ as big-endian (more significant bits will have smaller indices).

@@

&&&& $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \begin{cases}

@@

\big(\boldsymbol{\sigma}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\

@@

&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} - \hyperlink{L_but_64}{L}(\boldsymbol{\mu}_{\mathrm{g}}) + g'$ \\

@@

  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\

@@

&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A) + g'$ \\

@@

&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
  & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\
  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\

@@

&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
  & \text{if} \quad I_{\mathrm{e}} < 1024 \\
  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
